---
title: "Explorative Data Analysis of DCs infected with Yersinia scRNA-Seq data"
author: "Health Data Unit"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
  html_document:
    code_folding: hide
---


This report provides documentation for the analysis of scRNA-Seq data of Dendritic Cells 
progenitors infected with *Yersinia enterocolitica* provided by the Department of 
Internal Medicine II, University of Tübingen, Germany.


## Data Analysis

A repository storing custom scripts has been created and is available in github 
[here](https://github.com/caramirezal/interaction_DC-pathogens). This repository
along with the data are stored locally in the BioQuant curry cluster of the 
University of Heidelberg.

Data were preprocessed as shown in the `reports/scPipe_QC_workflow.Rmd` file.

```{r setup, include=FALSE}
## Dependencies
library(dplyr)
library(scPipe)
library(ggplot2)
library(rmarkdown)
library(Seurat)
library(ggrepel)

knitr::opts_chunk$set(
        warning = FALSE, 
        message = FALSE,
        cache = FALSE,
        results = 'hide',
        fig.align='center'
)

## Setting paths
path2project <- '/media/ag-cherrmann/cramirez/interaction_DC-pathogens/'
knitr::opts_knit$set(
        root.dir = path2project
)
```



```{r, settings}
## Loading data
seurat <- readRDS('data/yersinia_seurat.rds')

## Reading processed annotations
anns <- readxl::read_xlsx('data/20201214 C025_sample_sheet_SEA.xlsx') %>%
            as.data.frame()
rownames(anns) <- with(anns, paste0(plate_number, '_', well_position))
## Adding annotations
anns <- anns[rownames(seurat@meta.data), ]
anns <- anns[, c("progenitor_type: 'CDP' or 'MDP'",
                 "bacteria: 'HK Ye' or 'Ye' or 'no Ye'",
                 "treatment: A-untreated; B-MOI 1; C-MOI 0.1; D-MOI 0.01",
                 "bacteria_attached: '+' or '-'",
                 'infection_time')]
colnames(anns)[1:4] <- c('progenitor', 'bacteria', 'treatment', 'bacteria_attached')
## row order check
any(rownames(anns) != rownames(seurat@meta.data))
seurat@meta.data <- cbind(seurat@meta.data, anns)

## Selected category
category <- 'plate_number'
```

```{r}
seurat@meta.data %>%
        paged_table()
```


## Filtering cells

Initially, there were `r ncol(seurat)` wells containing single cells.
To filter out poor quality sequenced wells we removed some of them as follows: 
*i*) cells above 95 percentile for the value of the percentage of gene counts
corresponding to mitochondrial genes and *ii*) cells under and above 5 and 95 
percentile threshold of number of RNA counts.


```{r, filtering}
seurat@meta.data %>%
  ggplot(aes(x=log(nCount_RNA), 
             colour=get(category),
             y=percent.mt)) +
        geom_point(alpha=0.5) +
        geom_hline(yintercept = quantile(seurat$percent.mt, 0.95),
                   linetype = 'dashed', color='red') +
        geom_vline(xintercept = quantile(log(seurat$nCount_RNA), c(0.05, 0.95)),
                   linetype = 'dashed', color='red') +
        labs(x='Log(Counts of RNA)', y='Percentage mitochondrial',
             colour='') +
        theme_classic()
```

```{r, filtering_cellss}
seurat <- subset(seurat,
                 quantile(seurat$percent.mt, 0.05) < percent.mt & 
                     percent.mt < quantile(seurat$percent.mt, 0.95) &
                         nCount_RNA < quantile(seurat$nCount_RNA, 0.95)
)
```

After filtering `r ncol(seurat)` wells passed the quality control. 

# Single cells

The next table shows the number of wells for single cells or pools in each plate after
filtering. There are `r 179 + 14 + 157` single cells across three different plates (P193, P194 and P196). In the following sections we focused on the analysis of single cells.


```{r, single_cells_table, results='asis'}
table(seurat$sample_name, seurat$plate_number) %>%
        as.data.frame.matrix() %>%
        paged_table()
```


```{r, single_cells_umap, fig.align='center'}
seurat.sc <- subset(seurat, sample_name == 'single cell')
seurat.sc <- NormalizeData(seurat.sc) %>% ScaleData()
seurat.sc <- FindVariableFeatures(seurat.sc)
seurat.sc <- RunPCA(seurat.sc, npcs = 50, features = VariableFeatures(seurat.sc))
seurat.sc <- RunUMAP(seurat.sc, 
                     reduction = 'pca', 
                     dims = 1:20)
```



The next UMAP projections show single cells according to the different categories 
in clusters. Four clusters can be observed which coincides with cells in plates. 
Some cells of plate 196 (blue) fall more close to that of plates 193 (salmon) and 194 (green). This suggest the existence of a batch effect ought to differences in 
sequencing of different plates. 


```{r categories_in_single_cells, fig.align='center', fig.width=12, fig.height=11}
## Annotating with yersinia infection and cell type
plate <- seurat.sc %>%
        DimPlot(label = TRUE, 
                label.box = TRUE, 
                repel = TRUE) + NoLegend() +
        ggtitle('Plate')

disease.state <- seurat.sc %>%
        DimPlot(group.by = c('disease_state'),
                label = TRUE,
                label.box = TRUE) + 
        NoLegend() +
        ggtitle('Disease State')

bacteria <- seurat.sc %>%
        DimPlot(group.by = c('bacteria'),
                label = TRUE,
                label.box = TRUE, 
                repel = TRUE) + 
        NoLegend() +
        ggtitle('Bacteria')

treatment <- seurat.sc %>%
        DimPlot(group.by = c('treatment'),
                label = TRUE,
                label.box = TRUE, 
                repel = TRUE) + 
        NoLegend() +
        ggtitle('Treatment')

progenitor <- seurat.sc %>%
        DimPlot(group.by = c('progenitor'),
                label = TRUE,
                label.box = TRUE, 
                repel = TRUE) + 
        NoLegend() +
        ggtitle('Progenitor')

infection_time <- seurat.sc %>%
        DimPlot(group.by = c('infection_time'),
                label = TRUE,
                label.box = TRUE, 
                repel = TRUE) + 
        NoLegend() +
        ggtitle('Infection time')

gridExtra::grid.arrange(
  grobs = list(plate, disease.state, 
               bacteria, treatment, 
               progenitor, infection_time)
)
```

# Differential expression analysis

We performed a differential expression analysis taking cells in plates separately.

## 1 HPI

First we analyzed 1 hour post infection (HPI) which corresponds to plate 193. The 
next table shows the composition of the plate 193 according to the progenitor cell
types and whether the bacteria was found to be attached or not.

```{r}
seurat.sc@meta.data %>%
  filter(plate_number=='P193') %>%
  select(bacteria, progenitor) %>%
  table()
```

### CDP 

 * Infected vs non-infected 
 
The next table shows DEGs in Dendritic cells derived from CDP comparing infected *vs*
non-infected cells.


```{r, CDP_infection_1hpi}
Idents(seurat.sc) <- seurat.sc$bacteria
cdp.1hpi.degs <- seurat.sc %>%
  subset(progenitor=='CDP' & plate_number=='P193') %>%
  FindMarkers(ident.1='Ye') %>%
  arrange(desc(abs(avg_log2FC))) %>%
  filter(p_val < 0.05)
cdp.1hpi.degs %>%
  paged_table()
```

```{r}
seurat.sc %>%
  subset(progenitor=='CDP' & plate_number=='P193') %>%
  FindMarkers(ident.1='Ye') %>%
  arrange(desc(abs(avg_log2FC))) %>%
  add_rownames('gene') %>%
  mutate(highlight=ifelse( ( abs(avg_log2FC) > 0.55  & p_val < 0.05) | 
                            p_val < 0.001, TRUE, FALSE)) %>%
  mutate(label=ifelse(highlight==TRUE, gene, '')) %>%
  ggplot(aes(x=avg_log2FC, y=-log10(p_val),
             colour=highlight,
             label=label)) +
        geom_point(show.legend = FALSE) +
        geom_text_repel(show.legend=FALSE) +
        theme(legend.position = 'none') +
        scale_color_manual(values = c('salmon', 'black')) +
        theme_classic()
```


### MDP 

 * Infected vs non-infected 
 
The next table shows DEGs in Dendritic cells derived from MDP comparing infected *vs*
non-infected cells.


```{r, CDP_infection}
Idents(seurat.sc) <- seurat.sc$bacteria
mdp.1hpi.degs <- seurat.sc %>%
  subset(progenitor=='MDP' & plate_number=='P193') %>%
  FindMarkers(ident.1='Ye') %>%
  arrange(desc(abs(avg_log2FC))) %>%
  filter(p_val < 0.05)
mdp.1hpi.degs %>%
  paged_table()
```

```{r}
seurat.sc %>%
  subset(progenitor=='MDP' & plate_number=='P193') %>%
  FindMarkers(ident.1='Ye') %>%
  arrange(desc(abs(avg_log2FC))) %>%
  add_rownames('gene') %>%
  mutate(highlight=ifelse( ( abs(avg_log2FC) > 0.8  & p_val < 0.05 ) |
                             p_val < 0.0005,
                           TRUE, FALSE)) %>%
  mutate(label=ifelse(highlight==TRUE, gene, '')) %>%
  ggplot(aes(x=avg_log2FC, y=-log10(p_val),
             colour=highlight,
             label=label)) +
        geom_point(show.legend = FALSE) +
        geom_text_repel(show.legend=FALSE) +
        xlim(c(-2, 2)) +
        theme(legend.position = 'none') +
        scale_color_manual(values = c('salmon', 'black')) +
        theme_classic()
```


## 24 HPI

Then, we analyzed 24 hour post infection (HPI) which corresponds to plate 196. The 
next table shows the composition of the plate 196 according to the progenitor cell
types and whether the bacteria was found to be attached or not.


```{r plate_196}
seurat.sc@meta.data %>%
  filter(plate_number=='P196') %>%
  select(bacteria, progenitor) %>%
  table()
```


### CDP 

 * Infected vs non-infected 
 
The next table shows DEGs in Dendritic cells derived from CDP comparing infected *vs*
non-infected cells.


```{r, CDP_infection_24hpi}
Idents(seurat.sc) <- seurat.sc$bacteria
cdp.24hpi.degs <- seurat.sc %>%
  subset(progenitor=='CDP' & plate_number=='P196') %>%
  FindMarkers(ident.1='Ye') %>%
  arrange(desc(abs(avg_log2FC))) %>%
  filter(p_val < 0.05)
cdp.24hpi.degs %>%
  paged_table()
```



```{r}
seurat.sc %>%
  subset(progenitor=='CDP' & plate_number == 'P196') %>%
  FindMarkers(ident.1='Ye') %>%
  arrange(desc(abs(avg_log2FC))) %>%
  add_rownames('gene') %>%
  mutate(highlight=ifelse(abs(avg_log2FC) > 1.25 & 
                            p_val < 0.05, TRUE, FALSE)) %>%
  mutate(label=ifelse(highlight==TRUE, gene, '')) %>%
  ggplot(aes(x=avg_log2FC, y=-log10(p_val),
             colour=highlight,
             label=label)) +
        geom_point(show.legend = FALSE) +
        geom_text_repel(show.legend=FALSE, force = 2) +
        theme(legend.position = 'none') +
        scale_color_manual(values = c('salmon', 'black')) +
        theme_classic()
```


### MDP 

 * Infected vs non-infected 
 
The next table shows DEGs in Dendritic cells derived from MDP comparing infected *vs*
non-infected cells.


```{r, MDP_infection_24hpi}
Idents(seurat.sc) <- seurat.sc$bacteria
mdp.24hpi.degs <- seurat.sc %>%
  subset(progenitor=='MDP' & plate_number=='P196') %>%
  FindMarkers(ident.1='Ye') %>%
  arrange(desc(abs(avg_log2FC))) %>%
  filter(p_val < 0.05)
mdp.24hpi.degs %>%
  paged_table()
```


```{r}
seurat.sc %>%
  subset(progenitor=='MDP' & plate_number == 'P196') %>%
  FindMarkers(ident.1='Ye') %>%
  arrange(desc(abs(avg_log2FC))) %>%
  add_rownames('gene') %>%
  mutate(highlight=ifelse(abs(avg_log2FC) > 1.25 & 
                            p_val < 0.05, TRUE, FALSE)) %>%
  mutate(label=ifelse(highlight==TRUE, gene, '')) %>%
  ggplot(aes(x=avg_log2FC, y=-log10(p_val),
             colour=highlight,
             label=label)) +
        geom_point(show.legend = FALSE) +
        geom_text_repel(show.legend=FALSE, force = 2) +
        theme(legend.position = 'none') +
        scale_color_manual(values = c('salmon', 'black')) +
        theme_classic()
```