---
title: "Analysis of pooled cells"
author: "Health Data Unit"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    code_folding: hide
---


This R markdown provides documentation about the analysis of pooled Dendritic Cells (DCs) infected
with *Yersinia enterocolitica* provided by the Department of Internal Medicine II, University of 
Tübingen, Germany.

From the scRNA-Seq analysis the following observations were done:

*i)* Bystander cells cluster together with mock-infected cells in MDP, but with infected cells
in MDP at 1 hpi.

*ii)* Spi1 is more active in infected CDP cells at 1 hpi.

*iii)* Jun is more active in bystander cells in CDPs at 1 hpi.

*iv)* DEA reveals different gene expression patterns in infected, bystander and mock-infected
cells.


In order to validate these observations we further analyze RNA-Seq data from pooled
cells.


```{r setup, include=FALSE}
library(ggplot2)
library(dplyr)
library(rmarkdown)
library(reshape2)
library(Seurat)
library(ggrepel)
library(pheatmap)
library(viridis)
library(gridExtra)

set.seed(333)

knitr::opts_chunk$set(
        warning = FALSE, 
        message = FALSE,
        cache = FALSE,
        fig.align='center'
)

## Setting paths
path2project <- '/media/ag-cherrmann/cramirez/interaction_DC-pathogens/'
knitr::opts_knit$set(
        root.dir = path2project
)
```



```{r, settings}
## Loading data
seurat <- readRDS('data/yersinia_seurat.rds')

## Reading processed annotations
anns <- readxl::read_xlsx('data/20201214 C025_sample_sheet_SEA.xlsx') %>%
            as.data.frame()
rownames(anns) <- with(anns, paste0(plate_number, '_', well_position))
## Adding annotations
anns <- anns[rownames(seurat@meta.data), ]
anns <- anns[, c("progenitor_type: 'CDP' or 'MDP'",
                 "bacteria: 'HK Ye' or 'Ye' or 'no Ye'",
                 "treatment: A-untreated; B-MOI 1; C-MOI 0.1; D-MOI 0.01",
                 "bacteria_attached: '+' or '-'",
                 'infection_time')]
colnames(anns)[1:4] <- c('progenitor', 'bacteria', 'treatment', 'bacteria_attached')
## row order check
#any(rownames(anns) != rownames(seurat@meta.data))
seurat@meta.data <- cbind(seurat@meta.data, anns)

## Selected category
category <- 'plate_number'
```

## Selecting pooled cells

The next table show the content of the seurat object for pooled cells.

```{r, subsetting, results='asis'}
seurat <- subset(seurat, sample_name != 'single cell')
table(seurat$sample_name, seurat$plate_number) %>%
        as.data.frame.matrix() %>%
        paged_table()
```

It can be seen that most of the pooled cells types consist of 100 populations. Then, here we
focused in this sample type.


## QC metrics

The next violin plots show different QC metrics comparing the demultiplexed data analyzed here
versus a previous demultiplexed.


```{r, comparisson_old_new_metrics, fig.width=14}
new.qc <- seurat@meta.data %>%
  filter(sample_name == '100 cell popn') %>%
  mutate(plate=gsub('_.*', '', well)) %>%
    dplyr::select(plate, nCount_RNA, 
                  nFeature_RNA, 
                  unaligned:mapped_to_MT) %>%
    melt() %>%
    filter(variable!='mapped_to_ERCC') %>%
    ggplot(aes(x=plate, y=log10(value), fill=plate)) +
          geom_violin() +
          facet_wrap(~variable) +
          theme_bw() +
          theme(legend.position = 'none') +
          labs(x='', y='Log10(Counts)') +
          ggtitle('New demultiplexing')

old.demultiplexing <- readRDS('data/DCs-pathogen_interactions_seu.rds')
old.qc <- old.demultiplexing@meta.data %>%
  add_rownames(var = 'well') %>%
  filter(sample_name == '100 cell popn') %>%
    dplyr::select(plate_number, nCount_RNA, 
                  nFeature_RNA, 
                  unaligned:mapped_to_MT) %>%
    melt() %>%
    filter(variable!='mapped_to_ERCC') %>%
    ggplot(aes(x=plate_number, y=log10(value), fill=plate_number)) +
          geom_violin() +
          facet_wrap(~variable) +
          theme_bw() +
          theme(legend.position = 'none') +
          labs(x='', y='Log10(Counts)') +
          ggtitle('Old demultiplexing')

gridExtra::grid.arrange(
  new.qc,
  old.qc,
  ncol=2
)
```

Plate 196 seems to have a very low number of reads and therefore is filtered out. Additionally,
cells with less than 1000 UMI are also removed.


```{r, filtering}
counts_th <- 1000

## Number of RNA counts vs % mito
seurat@meta.data %>% 
  ggplot(aes(x=log(nCount_RNA), 
             colour=get(category),
             y=percent.mt)) +
        geom_point(alpha=0.5) +
        labs(x='Log(Counts of RNA)', y='Percentage mitochondrial',
             colour='') +
        geom_vline(xintercept = log(counts_th), 
                   linetype = 'dashed',
                   colour='red') +
        theme_classic()

#seurat <- subset(seurat, 
#                 nCount_RNA > counts_th & 
#                         plate_number != 'P196')
```

## UMAP projection

The following UMAP show the pooled cells according to categories.


```{r, umap, fig.align='center'}
seurat <- NormalizeData(seurat) %>% ScaleData()
seurat <- FindVariableFeatures(seurat)
seurat <- RunPCA(seurat, 
                    npcs = 50, 
                    features = VariableFeatures(seurat))
seurat <- RunUMAP(seurat, 
                     reduction = 'pca', 
                     dims = 1:20)
```


```{r, umap_by_plate}
## Annotating with yersinia infection and cell type
seurat$treatment[is.na(seurat$treatment)] <- 'NA'
plate <- seurat %>%
        DimPlot(repel = TRUE)  +
        ggtitle('Plate')
#plate + geom_vline(xintercept = -8, 
#                   linetype = 'dashed',
#                   colour='red')
plate
```


We can observe an outlier cluster of cells. We further removed these cells 
shown in the UMAP below the threshold UMAP_1 < -8.


```{r categories_in_single_cells, fig.align='center', fig.width=10, fig.height=16}


## Annotating with yersinia infection and cell type
seurat$treatment[is.na(seurat$treatment)] <- 'NA'
plate <- seurat %>%
        DimPlot(repel = TRUE)  +
        ggtitle('Plate')

disease.state <- seurat %>%
        DimPlot(group.by = 'disease_state') + 
        ggtitle('Disease State')

bacteria <- seurat %>%
        DimPlot(group.by = c('bacteria_attached'),
                label = TRUE,
                repel = TRUE) + 
        NoLegend() +
        ggtitle('Bacteria')

treatment <- seurat %>%
        DimPlot(group.by = c('treatment'),
                label = TRUE,
                repel = TRUE) + 
        NoLegend() +
        ggtitle('Treatment')

progenitor <- seurat %>%
        DimPlot(group.by = c('progenitor'),
                label = TRUE,
                repel = TRUE) + 
        NoLegend() +
        ggtitle('Progenitor')

infection_time <- seurat %>%
        DimPlot(group.by = c('infection_time')) + 
        ggtitle('Infection time')

sample_name <- seurat %>%
        DimPlot(group.by = c('sample_name')) + 
        ggtitle('Sample type')

gridExtra::grid.arrange(
  grobs = list(plate, disease.state, 
               bacteria, treatment, 
               progenitor, infection_time,
               sample_name),
  ncol = 2
)
```


## Definition of the bystander cells

The next table show the definition of bystander cells according to the treatment and
infection status per plate and progenitor type.


```{r}
#seurat <- subset(seurat, UMAP_1 > -8)
seurat$'treated' <- ifelse(seurat$treatment == 'MOI 0', 
                              'Mock-Infected', 'Infected')
seurat$'Bystander' <- paste0(seurat$treated, '_', seurat$bacteria_attached) 
seurat$Bystander <- gsub(' ', '_', seurat$Bystander)
seurat$Bystander <- plyr::mapvalues(
  seurat$Bystander,
  from = c('Infected_-', 'Infected_+', 'Mock-Infected_-'),
  to = c('Bystander', 'Infected', 'Mock-Infected')
)

seurat@meta.data %>%
        filter(plate_number=='P193') %>%
  dplyr::select(treatment, bacteria_attached, 
                progenitor, Bystander) %>%
  table() %>%
  as.data.frame() %>%
  filter(Freq > 0) %>%
  arrange(treatment) %>%
  paged_table()
```


## 1 HPI (Plate 193) analysis

The next UMAP show only plate 193 across experimental conditions.


```{r, results='hide'}
## Subsetting to plate 193
seurat.p193 <- subset(seurat, plate_number=='P193')

## normalization
seurat.p193 <- NormalizeData(seurat.p193) %>% 
                  ScaleData() %>%
                        RunPCA(npcs = 30) %>%
                        RunUMAP(dims=1:20)


disease.state.p193 <- seurat.p193 %>%
        DimPlot(group.by = 'disease_state',
                label = TRUE,
                repel = TRUE) +  
        NoLegend() +
        ggtitle('Disease State')

bacteria.p193 <- seurat.p193 %>%
        DimPlot(group.by = c('bacteria_attached'),
                label = TRUE,
                repel = TRUE) + 
        ggtitle('Bacteria')

treatment.p193 <- seurat.p193 %>%
        DimPlot(group.by = c('treatment')) +
        ggtitle('Treatment')

progenitor.p193 <- seurat.p193 %>%
        DimPlot(group.by = c('progenitor')) 
        ggtitle('Progenitor')

gridExtra::grid.arrange(
  grobs = list(disease.state.p193, 
               bacteria.p193, 
               treatment.p193, 
               progenitor.p193),
  ncol = 2
)

```



## Gene expression correlation

The next heatmap shows spearman correlation between all CDP pooled cells at time 1 hpi.
The diagonal values (=1) are set to zero. It can be seen that bystander cells cluster
together with mock-infected cells similar to what we observed in the single cells. 


### CDP

```{r, gene_expression_correlation_cdp_heatmap}
######################################################################################
## CDPs
p193.cdp <- subset(seurat.p193, progenitor=='CDP')
anns <- p193.cdp@meta.data %>% 
        select(Bystander)

## Correlation
p193.cdp.mtx <- p193.cdp@assays$RNA@scale.data[VariableFeatures(seurat.p193), ] %>% 
        cor(method = 'spearman')
diag(p193.cdp.mtx) <- 0

## Heatmap
p193.cdp.heatmap <- pheatmap(p193.cdp.mtx, 
         annotation_col = anns,
         annotation_colors = list(Bystander=c(`Infected`='red',
                                              `Bystander`='orange',
                                              `Mock-Infected`='green')), 
         show_rownames = FALSE,
         show_colnames = FALSE, 
         color = viridis(20)
)
p193.cdp.heatmap
```


### MDPs

Unfortunately, for MDPs there are not enough infected cells and the heatmap is not informative
as shown next.



```{r, , gene_expression_correlation_mdp_heatmap}
#################################################################################
## MDPs
p193.mdp <- subset(seurat.p193, progenitor=='MDP')
anns <- p193.mdp@meta.data %>% 
        select(Bystander)

## Correlation
p193.mdp.mtx <- p193.mdp@assays$RNA@scale.data[VariableFeatures(seurat.p193), ] %>% 
        cor(method = 'spearman')
diag(p193.mdp.mtx) <- 0

## Heatmap
p193.mdp.heatmap <- pheatmap(p193.mdp.mtx, 
         annotation_col = anns,
         annotation_colors = list(Bystander=c(`Infected`='red',
                                              `Bystander`='orange',
                                              `Mock-Infected`='green')), 
         show_rownames = FALSE,
         show_colnames = FALSE, 
         color = viridis(20)
)
p193.mdp.heatmap
```


## Spi1 regulon activity in CDP (1 hpi)

To validate elevated Spi1 activity in infected CDPs (ii) using the pool cells samples
we extracted the genes in the regulons inferred using pySCENIC. We selected the regulon
associated with the motif with higher score, and then ranked cells using the genes in 
that regulon with the `AddModuleScore` function from Seurat.


```{r, scoring_regulons, fig.width=12, fig.height=6}
## function to extract TFs in column 7 from SCENIC results
extract_targets <- function(x){
        targets <- regmatches(x, gregexpr("'[^']*'", x))[[1]] 
        gsub("'", '', targets)
}

## Reading spi1 regulons
regulons <- read.csv('analysis/scenic/reg.csv', skip = 2)

## Using the highest spi1 motif
spi1.regulon.df.top <- filter(regulons, TF == 'Spi1') %>%
                    arrange(desc(X)) %>% 
                       head(1)
spi1.regulon.top <- extract_targets(spi1.regulon.df.top$X.7)

p193 <- subset(seurat, plate_number=='P193')
p193 <- AddModuleScore(p193, 
                       features = list(spi1.regulon.top), 
                       name = 'spi1_regulon', 
                       nbin = 3, 
                       ctrl = 10, 
                       seed = 333)
p193@meta.data$Bystander <- factor(p193@meta.data$Bystander,
                                   levels = c('Mock-Infected',
                                              'Bystander',
                                              'Infected'))
spi1.plot.1h <- p193@meta.data %>%
        ggplot(aes(x=Bystander, 
                   y=log(spi1_regulon1),
                   fill=Bystander)) +
                geom_boxplot() +
                geom_jitter() +
                facet_wrap(~progenitor, 
                           scales = 'free') +
        theme_bw() +
        labs(x='', y='Log(Spi1 TFA Score)') +
        theme(legend.position = 'none') +
        ggtitle('Spi1 TFA') +
        scale_fill_manual(values = c('green',
                                      'orange',
                                      'red'))

## Using the highest spi1 motif
jun.regulon.df.top <- filter(regulons, TF == 'Jun') %>%
                    arrange(desc(X)) %>% 
                       head(1)
jun.regulon.top <- extract_targets(jun.regulon.df.top$X.7)

#p193 <- subset(seurat, plate_number=='P193')
p193 <- AddModuleScore(p193, 
                       features = list(jun.regulon.top), 
                       name = 'jun_regulon', 
                       nbin = 3, 
                       ctrl = 10, 
                       seed = 333)

p193@meta.data$Bystander <- factor(p193@meta.data$Bystander,
                                   levels = c('Mock-Infected',
                                              'Bystander',
                                              'Infected'))

jun.plot.1h <- p193@meta.data %>%
        ggplot(aes(x=Bystander, 
                   y=log(jun_regulon1),
                   fill=Bystander)) +
                geom_boxplot() +
                geom_jitter() +
                facet_wrap(~progenitor, 
                           scales = 'free') +
        theme_bw() +
        labs(x='', y='Log(Jun TFA Score)') +
        theme(legend.position = 'none') +
        ggtitle('Jun TFA') +
        scale_fill_manual(values = c('green',
                                      'orange',
                                      'red'))

gridExtra::grid.arrange(
  spi1.plot.1h,
  jun.plot.1h, 
  ncol=2
)
```



## 24 HPI analysis

The next UMAP show for 24 hpi pooled cells across experimental conditions.


```{r, results='hide', fig.height=16, fig.width=12}
## Modify here
seurat.24 <- subset(seurat, 
                    disease_state %in% c('Yesinia 24h post-infection',
                                       'Yesinia 22h post-infection'))

#seurat.p193 <- subset(seurat, plate_number=='P193')

## normalization
seurat.24 <- NormalizeData(seurat.24) %>% 
                  ScaleData() %>%
                        RunPCA(npcs = 30) %>%
                        RunUMAP(dims=1:20)


disease.state.24 <- seurat.24 %>%
        DimPlot(group.by = 'disease_state', pt.size = 3) + 
        theme(legend.position = 'top') +
        ggtitle('Disease State')

bacteria.24 <- seurat.24 %>%
        DimPlot(group.by = 'bacteria_attached', 
                pt.size = 3,
                label = TRUE,
                repel = TRUE) + 
        ggtitle('Bacteria')

treatment.24 <- seurat.24 %>%
        DimPlot(group.by = 'treatment', 
                pt.size = 3) +
        ggtitle('Treatment')

progenitor.24 <- seurat.24 %>%
        DimPlot(group.by = 'progenitor', 
                pt.size = 3) +
        ggtitle('Progenitor')
        
sample_type.24 <- seurat.24 %>%
        DimPlot(group.by = 'Bystander', 
                pt.size = 3,
                cols = c('orange', 'red', 'green')) +
        ggtitle('Infection status')
        
plate.24 <- seurat.24 %>%
        DimPlot(group.by = 'plate_number', 
                pt.size = 3) +
        ggtitle('Plate')

gridExtra::grid.arrange(
  grobs = list(disease.state.24, 
               bacteria.24, 
               treatment.24, 
               progenitor.24,
               sample_type.24,
               plate.24),
  ncol = 2
)

```

## Correlation analysis (24 hpi)




* CDP 

```{r, scores_24_only}
#################################################################################
## CDPs
cdp.24 <- subset(seurat.24, progenitor == 'CDP')
anns <- cdp.24@meta.data %>% 
        select(Bystander, 
               disease_state)

## Correlation
cdp.24.mtx <- cdp.24@assays$RNA@scale.data[VariableFeatures(seurat.p193), ] %>% 
        cor(method = 'spearman')
diag(cdp.24.mtx) <- 0

## Heatmap
cdp.24.heatmap <- pheatmap(cdp.24.mtx, 
         annotation_col = anns,
         annotation_colors = list(Bystander=c(`Infected`='red',
                                              `Bystander`='orange',
                                              `Mock-Infected`='green'),
                                  disease_state=c(`Yesinia 22h post-infection`='salmon',
                                                  `Yesinia 24h post-infection`='steelblue')), 
         show_rownames = FALSE,
         show_colnames = FALSE, 
         color = viridis(20)
)
cdp.24.heatmap
```

Two main clusters can be observed in the heatmap which can be associated to the
time points, being 22 and 24 hpi concentrated at the right and left, respectively.
It can also be appreciated that most of the infected cells are present at the
cluster corresponding to 24 hpi, which is in line with more activation at later 
time points. We also see that 22 hpi bystander cells cluster together with 
Mock-infected cells which agrees with the results observed using single cells.

### Using all cells (22 and 24 hpi)

```{r, scoring_regulons_boxplots_all, fig.width=12, fig.height=6}
seurat.24.all <- AddModuleScore(seurat.24, 
                       features = list(spi1.regulon.top), 
                       name = 'spi1_regulon', 
                       nbin = 3, 
                       ctrl = 10, 
                       seed = 333)
seurat.24.all@meta.data$Bystander <- factor(seurat.24.all@meta.data$Bystander,
                                   levels = c('Mock-Infected',
                                              'Bystander',
                                              'Infected'))
spi1.plot.all <- seurat.24.all@meta.data %>%
        ggplot(aes(x=Bystander, 
                   y=log(spi1_regulon1),
                   fill=Bystander)) +
                geom_boxplot() +
                geom_jitter() +
                facet_wrap(~progenitor, 
                           scales = 'free') +
        theme_bw() +
        labs(x='', y='Log(Spi1 TFA Score)') +
        theme(legend.position = 'none') +
        ggtitle('Spi1 TFA') +
        scale_fill_manual(values = c('green',
                                      'orange',
                                      'red'))

## Using the highest spi1 motif
jun.regulon.df.top <- filter(regulons, TF == 'Jun') %>%
                    arrange(desc(X)) %>% 
                       head(1)
jun.regulon.top <- extract_targets(jun.regulon.df.top$X.7)

#seurat.24.all <- subset(seurat, plate_number=='seurat.24.all')
seurat.24.all <- AddModuleScore(seurat.24.all, 
                       features = list(jun.regulon.top), 
                       name = 'jun_regulon', 
                       nbin = 3, 
                       ctrl = 10, 
                       seed = 333)

seurat.24.all@meta.data$Bystander <- factor(seurat.24.all@meta.data$Bystander,
                                   levels = c('Mock-Infected',
                                              'Bystander',
                                              'Infected'))

jun.plot.all <- seurat.24.all@meta.data %>%
        ggplot(aes(x=Bystander, 
                   y=log(jun_regulon1),
                   fill=Bystander)) +
                geom_boxplot() +
                geom_jitter() +
                facet_wrap(~progenitor, 
                           scales = 'free') +
        theme_bw() +
        labs(x='', y='Log(Jun TFA Score)') +
        theme(legend.position = 'none') +
        ggtitle('Jun TFA') +
        scale_fill_manual(values = c('green',
                                      'orange',
                                      'red'))

gridExtra::grid.arrange(
  spi1.plot.all,
  jun.plot.all, 
  ncol=2
)
```


## Scoring cells 24 hpi


```{r, scoring_regulons_24hpi, fig.width=12, fig.height=6}
seurat.24 <- subset(seurat.24, disease_state == 'Yesinia 24h post-infection')
seurat.24 <- AddModuleScore(seurat.24, 
                       features = list(spi1.regulon.top), 
                       name = 'spi1_regulon', 
                       nbin = 3, 
                       ctrl = 10, 
                       seed = 333)
seurat.24@meta.data$Bystander <- factor(seurat.24@meta.data$Bystander,
                                   levels = c('Mock-Infected',
                                              'Bystander',
                                              'Infected'))
spi1.plot.24h <- seurat.24@meta.data %>%
        ggplot(aes(x=Bystander, 
                   y=log(spi1_regulon1),
                   fill=Bystander)) +
                geom_boxplot() +
                geom_jitter() +
                facet_wrap(~progenitor, 
                           scales = 'free') +
        theme_bw() +
        labs(x='', y='Log(Spi1 TFA Score)') +
        theme(legend.position = 'none') +
        ggtitle('Spi1 TFA') +
        scale_fill_manual(values = c('green',
                                      'orange',
                                      'red'))

## Using the highest spi1 motif
jun.regulon.df.top <- filter(regulons, TF == 'Jun') %>%
                    arrange(desc(X)) %>% 
                       head(1)
jun.regulon.top <- extract_targets(jun.regulon.df.top$X.7)

#seurat.24 <- subset(seurat, plate_number=='seurat.24')
seurat.24 <- AddModuleScore(seurat.24, 
                       features = list(jun.regulon.top), 
                       name = 'jun_regulon', 
                       nbin = 3, 
                       ctrl = 10, 
                       seed = 333)

seurat.24@meta.data$Bystander <- factor(seurat.24@meta.data$Bystander,
                                   levels = c('Mock-Infected',
                                              'Bystander',
                                              'Infected'))

jun.plot.24h <- seurat.24@meta.data %>%
        ggplot(aes(x=Bystander, 
                   y=log(jun_regulon1),
                   fill=Bystander)) +
                geom_boxplot() +
                geom_jitter() +
                facet_wrap(~progenitor, 
                           scales = 'free') +
        theme_bw() +
        labs(x='', y='Log(Jun TFA Score)') +
        theme(legend.position = 'none') +
        ggtitle('Jun TFA') +
        scale_fill_manual(values = c('green',
                                      'orange',
                                      'red'))

gridExtra::grid.arrange(
  spi1.plot.24h,
  jun.plot.24h, 
  ncol=2
)
```

```{r, scoring_regulons_boxplots, fig.width=12, fig.height=6}

seurat.24 <- AddModuleScore(seurat.24, 
                       features = list(spi1.regulon.top), 
                       name = 'spi1_regulon', 
                       nbin = 3, 
                       ctrl = 10, 
                       seed = 333)
seurat.24@meta.data$Bystander <- factor(seurat.24@meta.data$Bystander,
                                   levels = c('Mock-Infected',
                                              'Bystander',
                                              'Infected'))
spi1.plot.24h <- seurat.24@meta.data %>%
        ggplot(aes(x=Bystander, 
                   y=log(spi1_regulon1),
                   fill=Bystander)) +
                geom_boxplot() +
                geom_jitter() +
                facet_wrap(~progenitor, 
                           scales = 'free') +
        theme_bw() +
        labs(x='', y='Log(Spi1 TFA Score)') +
        theme(legend.position = 'none') +
        ggtitle('Spi1 TFA') +
        scale_fill_manual(values = c('green',
                                      'orange',
                                      'red'))

## Using the highest spi1 motif
jun.regulon.df.top <- filter(regulons, TF == 'Jun') %>%
                    arrange(desc(X)) %>% 
                       head(1)
jun.regulon.top <- extract_targets(jun.regulon.df.top$X.7)

#seurat.24 <- subset(seurat, plate_number=='seurat.24')
seurat.24 <- AddModuleScore(seurat.24, 
                       features = list(jun.regulon.top), 
                       name = 'jun_regulon', 
                       nbin = 3, 
                       ctrl = 10, 
                       seed = 333)

seurat.24@meta.data$Bystander <- factor(seurat.24@meta.data$Bystander,
                                   levels = c('Mock-Infected',
                                              'Bystander',
                                              'Infected'))

jun.plot.24h <- seurat.24@meta.data %>%
        ggplot(aes(x=Bystander, 
                   y=log(jun_regulon1),
                   fill=Bystander)) +
                geom_boxplot() +
                geom_jitter() +
                facet_wrap(~progenitor, 
                           scales = 'free') +
        theme_bw() +
        labs(x='', y='Log(Jun TFA Score)') +
        theme(legend.position = 'none') +
        ggtitle('Jun TFA') +
        scale_fill_manual(values = c('green',
                                      'orange',
                                      'red'))

gridExtra::grid.arrange(
  spi1.plot.24h,
  jun.plot.24h, 
  ncol=2
)
```


## Scores all time points

```{r, scores_1_n_24_hpi, fig.width=12, fig.height=8}
gridExtra::grid.arrange(
  spi1.plot.1h,
  jun.plot.1h,
  spi1.plot.24h,
  jun.plot.24h,
  ncol=2
)
```


## Differential Expression Analysis

DEA across infected, bystander and mock-infected cells in CDPs at 1 hpi.

* CDP

```{r, plate_193_composition}
seurat@meta.data %>%
  filter(plate_number=='P193') %>%
  dplyr::select(Bystander, progenitor) %>%
  table() %>%
  as.data.frame() %>%
  filter(Freq>0) %>%
  paged_table()
```

```{r, degs_1hpi_cdp}
## Loading selected genes
selected_genes <- readLines('data/selected_genes.txt')

Idents(seurat) <- as.character(seurat$Bystander)
infvsBys.cdp.deg <- seurat %>%
  subset(progenitor=='CDP' & plate_number=='P193') %>%
  subset(Bystander != 'Mock-Infected') %>%
  FindMarkers(ident.1='Infected',
              logfc.threshold = 0,
              min.pct = 0,
              min.cells.feature = 0,
              min.cells.group = 0) %>%
  arrange(desc(abs(avg_log2FC))) %>%
  add_rownames('gene') %>%
  mutate(selected_gene=(toupper(gene) %in% toupper(selected_genes))) %>%
  filter(!grepl('^Gm', gene)) %>%
  mutate(highlight=ifelse( ( abs(avg_log2FC) > 0.4  & p_val < 0.05 ),
                           TRUE, FALSE) ) %>%
  mutate(label=ifelse(highlight==TRUE | selected_gene ==TRUE, gene, ''))
infvsBys.cdp.plot <- infvsBys.cdp.deg %>%
  ggplot(aes(x=avg_log2FC, y=-log10(p_val),
             colour=highlight,
             label=label)) +
        geom_point(show.legend = FALSE) +
        geom_point(data=subset(infvsBys.cdp.deg,
                               selected_gene==TRUE), 
                   aes(colour='green'),
                   show.legend=FALSE) +
        geom_text_repel(show.legend=FALSE, 
                        colour='black') +
        theme(legend.position = 'none') +
        scale_color_manual(values = c('salmon', 'green', 
                                      'black', 'red')) +
        theme_classic() +
        xlim(-2, 2) +
        ggtitle('CDP - 1HPI - Infected vs Bystander')

```

```{r table_cdp_inf_vs_bys_1hpi}
infvsBys.cdp.deg %>%
  filter(p_val < 0.05) %>%
  paged_table()
```



