---
title: "Data preprocessing"
author: "Health Data Science Unit"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
  html_document:
    code_folding: hide
---

```{r setup, include=FALSE}
## Dependencies
library(UpSetR)
library(readxl)
library(knitr)
library(dplyr)

knitr::opts_chunk$set(
        warning = FALSE, 
        message = FALSE,
        cache = FALSE
)

## Setting paths
path2project <- '/media/ag-cherrmann/cramirez/interaction_DC-pathogens/'
knitr::opts_knit$set(
        root.dir = path2project
)
```

## Illumina indexes in fastq files

We first checked the distribution of Illumina indexes 
in the fastq files. 
Illumina indexes were extracted from the headers in the fastq files (R1)
and then we filtered out sequencing error by matching them with
the ones in the sample sheet (column = illumina_index_index_sequence_separate_index_read).
The next upset plot show the sets of filtered illumina indexes for each
fastq file. It can be seen that each fastq file contain a single index.
The upset plot shows also the comparisson to the indexes reported
in the sample sheet.

```{r, illumina_indexes_in_fastqs, eval=FALSE}
fastq.path <- paste0(path2project, 'data/fastq')
fastq.paths <- list.files(fastq.path, full.names = T)
fastq.paths <- grep('R1.fastq.gz', fastq.paths, value = T)

fastq.contents <- lapply(fastq.paths, readLines)
names(fastq.contents) <- gsub(path2project, '', fastq.paths)
fastq.contents <- lapply(fastq.contents, 
                         function(x) grep('@', x, value = T))
ill.idxs <- lapply(fastq.contents, function(x) substr(x, 49, 56))
ill.idxs <- lapply(ill.idxs, function(x) gsub('[0-9]|:', '', x))
ill.idxs <- lapply(ill.idxs, unique)
saveRDS(ill.idxs,
        'analysis/illum_idxs_in_fastqs.rds')
```
```{r fig.align='center'}
## Reading Illumina indexes from fastq files
ill.idxs <- readRDS('analysis/illum_idxs_in_fastqs.rds')
names(ill.idxs) <- gsub('.*\\/|.fastq.gz', '', names(ill.idxs))

## Reading annotations
metadata <- readRDS('data/sample_sheet.rds')
ill.idxs.ann <- unique(metadata$illumina_index_index_sequence_separate_index_read)
ill.idxs <- lapply(ill.idxs, function(x) x[x %in% ill.idxs.ann])
ids.sheet <- metadata$illumina_index_index_sequence_separate_index_read %>%
                unique()
ill.idxs$'idxs_ssheet' <- ids.sheet
#ill.idxs <- c(ill.idxs, idxs_ssheet=ids.sheet)
upset(fromList(ill.idxs), nsets = 100)
```

The next table show the Illumina indexes in each fastq file.

```{r fastq_content_table}
fastq.idxs <- data.frame(file=names(ill.idxs)[1:6],
                         index=unlist(ill.idxs[1:6]))
kable(fastq.idxs, row.names = FALSE)
```

